---
title: "Jessica Qin Final Project"
output:
  html_document:
    df_print: paged
---

#Gathering Data

The data is from the World Happiness Report which consists of a Positive Affect attribute (which represents happiness, laughter, and enjoyment), a Negative Affect attribute (which represents worry, sadness, and anger) as well as various other attributes relevant to the country's government and social aspects throughout various years.

I chose this dataset because although each person's happiness is dependent on their own situation, I wanted to see if there was anything that impacted the countries happiness as a whole.

_Attributes:_

Country

Year

Life Expectancy

Freedom: Average of Binary response (1 = satisfied or 0 = dissatisfied) to the question "Are you satisfied or dissatisfied with your freedom to choose what you do with your life"

Generosity: The residual of regressing national average of reponse to "Have you donated money to a charity in the past month?" on GDP per capita

Corruption: Average of Binary response (1 = yes or 0 = no) to the question "Is corruption widespread throughout the government or not" and "Is corruption widespread within businesses or not?"

Positive Affect: Average of binary answers to "Did you experience a lot of happiness today?" and "DId you smile or laugh a lot yesterday?"

Negative Affect: Average of binary answers to "Did you experience a lot of sadness today?" and "Did you experience a lot of these feelings yesterday?"


```{r}
library(tidyverse)

df <- read_csv("happiness.csv")


select_df <- df %>%
  select(`Country name`,Year,`Healthy life expectancy at birth`, `Freedom to make life choices`, `Perceptions of corruption`, Generosity, `Positive affect`, `Negative affect`) 

colnames(select_df) <- c("Country", "Year", "life_exp", "Freedom", "Corruption", "Generosity", "positive_affect", "negative_affect") 

select_df

```


#Tidying Data

There's a lot of missing data in this dataset for various reasons. There are a handful of countries that are missing values for Freedom and Corruption, this is usually for countries that are ruled under a dictatorship so the reason behind the missing data makes sense beause these countries most likely do not want it broadcasted that their citizens do not believe they have any freedom or that they feel the government is corrupted.

For some of the NA values, I replace them with the mean of the rest of the attribute values for that country. For countries that are missing all of their attributes for that column, I set the value to 0.5.


```{r}

#Replacing NA with mean in the Freedom Column
select_df <- select_df %>%
  group_by(Country)%>%
  mutate(Freedom = ifelse(is.na(Freedom), mean(Freedom, na.rm = TRUE), Freedom))

#Replacing NA with mean in the Generosity Column
select_df <- select_df %>%
  group_by(Country)%>%
  mutate(Generosity = ifelse(is.na(Generosity), mean(Generosity, na.rm = TRUE), Generosity))

#Replacing NA with mean in the Negative Affect Column
select_df <- select_df %>%
  group_by(Country)%>%
  mutate(positive_affect = ifelse(is.na(positive_affect), mean(positive_affect, na.rm = TRUE), positive_affect))

#Replacing NA with mean in the Negative Affect Column
select_df <- select_df %>%
  group_by(Country)%>%
  mutate(negative_affect = ifelse(is.na(negative_affect), mean(negative_affect, na.rm = TRUE), negative_affect))

#Replacing NA with mean in the Corruption Column
#If there is no corruption data for that country, I set it to 0.5
select_df <- select_df %>%
  group_by(Country)%>%
  mutate(Corruption = ifelse(is.na(Corruption), mean(Generosity, na.rm = TRUE), Corruption)) %>%
  mutate(Corruption = ifelse(Corruption < 0, 0.5, Corruption))

select_df
```

#Top 10 most powerful countries

According to https://www.usnews.com/news/best-countries/power-rankings 

```{r}
country_df <- select_df %>%
  filter(Year >= 2010) %>%
  filter( Country == "United States" | Country == "Russia" | Country == "China" | Country == "Germany" | Country == "United Kingdom" | Country == "France" | Country == "Japan" | Country == "Israel" | Country == "Saudi Arabia" | Country == "South Korea")

country_df
  
```

#Looking for Trends

Looking at the evolution of happiness for al countries in the dataset:

```{r}
country_df %>%
  ggplot(mapping = aes(x = Year, y = positive_affect, group_by(Year))) + 
  geom_smooth(size = 1, se = FALSE) +
  ylim(0.68,0.78) +
  ggtitle("Evolution of Global Happiness")

```

Overall, there is no general trend for global happiness and year, this is probably because while some countries are increasing in happiness rating, others are decrease.

Here I plot Positive Affect vs Years to see if there is a trend in the evolution of happiness for the top 10 most powerful countries:

```{r}
country_df %>%
  ggplot(mapping = aes(x = Year, y = positive_affect, group_by(Year), color = Country)) + 
  geom_smooth(size = 1, se = FALSE) +
    ggtitle("Evolution of Happiness top 10 most powerful countries")

```


There appears to be no common trend for the 10 most powerful countries, a lot of the very developed countries such as Germany, Japan, South Korea, and the United States have slighlty declined throughout the years. Its mainly the currently developing countries that experience an increase in happiness. 

Developed Countries:

```{r}
country_df %>%
  filter(Country == "France" | Country == "Germany" | Country == "Japan" | Country == "South Korea" | Country == "United Kingdom" | Country == "United States" ) %>%
  ggplot(mapping = aes(x = Year, y = positive_affect, group_by(Year), color = Country)) + 
  geom_smooth(size = 1, se = FALSE) +
    ggtitle("Evolution of Happiness top 10 most powerful countries(Developed)")


```

Developing Countries:


```{r}
country_df %>%
  filter(Country == "China" | Country == "Israel" | Country == "Russia" | Country == "Saudi Arabia" ) %>%
  ggplot(mapping = aes(x = Year, y = positive_affect, group_by(Year), color = Country)) + 
  geom_smooth(size = 1, se = FALSE) +
  ggtitle("Evolution of Happiness top 10 most powerful countries(Developing)")

```

#Some Interesting trends to look at

#Positive Affect vs Life Expectancy

Below I plotted the positive affect vs life expectancy to see if a country's life expectancy impacts the citizen's happiness

Top 10 powerful countries:

```{r}
country_df %>%
  ggplot(mapping = aes(x = life_exp, y = positive_affect)) + 
  geom_smooth(size = 1, se = FALSE) +
  ggtitle("Happiness vs Life Expectancy for the top 10 most powerful countries")
```

All of the countries:

```{r}
select_df %>%
  ggplot(mapping = aes(x = life_exp, y = positive_affect)) + 
  geom_smooth(size = 1, se = FALSE) +
  ggtitle("Happiness vs Life Expectancy")
```

I anticipated that the higher the life expectancy, the higher the level of happiness. but the happiness rating peaks at the life expectancy of 70 and declines after that which I found interesting. 


The relationship between Happiness and Lfife Expectancy is the same for the top 10 most powerful countries and that of the rest of the world

#Positive Affect vs Generosity

```{r}
select_df%>%
  ggplot(mapping = aes(x = Generosity, y = positive_affect)) + 
  geom_smooth(size = 1, se = FALSE) +
  ggtitle("Happiness vs Generosity")
```

As expected, the trend for this is the higher the generosity, higher the happiness.


#Positive Affect vs Corruption

```{r}
country_df%>%
  ggplot(mapping = aes(x = Corruption, y = positive_affect)) + 
  geom_smooth(size = 1, se = FALSE) +
  ggtitle("Happiness vs Government Corruption")

```

The trend appears the higher the corruption, lower the happiness rating.
What's interesting is that a little corruption appears to not impact people's happiness however when over 70% of the population sees the government as corrupt, there appears to be a strong, negative, linear relationship between corruption and happiness.

#Hypothesis Testing:

Let's take another look at the evolution of Happiness for the top 10 most powerful countries and see if we can fit a linear regression machine learning model to test if there is a relationship between year and happiness

```{r}
country_df %>%
  ggplot(mapping = aes(x = Year, y = positive_affect, group_by(Year), color = Country)) + 
  geom_smooth(size = 1, se = FALSE) +
  ggtitle("Evolution of Happiness for the top 10 most powerful countries")
```

Since we can see in the plot that each Country has a different trend, we will include an interaction term for Country and Year.

```{r}
library(ISLR)
library(tidyverse)
library(broom)

#Interaction term between year and country
mdl <- lm(positive_affect~Year+Country+Country:Year, data=country_df)

summary(mdl)

mdl_stats <- mdl %>%
  tidy()

mdl_stats
```


Now for each country we can look at their P values to see if we reject the null hypothesis of NO RELATIONSHIP between year and Happiness.

We reject the nul hypothesis if our p value is less than $\alpha = 0.05$

China: p = 0.0891710097 $\nless$ 0.05 So we DO NOT reject the hypothesis

France: p = 0.0812782971 $\nless$ 0.05 So we DO NOT reject the hypothesis

Germany: p = 0.0123077267 $\lt$ 0.05 so we reject

Israel: p = 0.0134298160 $\lt$ 0.05 so we reject

Japan: p = 0.0003156129 $\lt$ 0.05 so we reject

Russia: p = 0.1679674502 $\nless$ 0.05 so we DO NOT reject

Saudia Arabia: p = 0.1241861861 $\nless$ 0.05 so we DO NOT reject

South Korea: p = 0.0981772404 $\nless$ 0.05 so we DO NOT reject

United Kingdom: p = 0.0002117193 $\lt$ 0.05 so we reject

United States: p = 0.0235877448 $\lt$ 0.05 so we reject

#Conclusion

From the linear regression model we can identify which countries have a trend regarding their happiness rating. 
Germany, Israel, Japan, United Kingdom, and United States are decreasing.


For the Countries that do show a correlation, we can use out model to calculate an estimate for happiness changes per year


```{r}
# Extract the coefficients
mdl_m <- summary(mdl)$coefficients


yr <- mdl_m["Year","Estimate"]
germany <- mdl_m["Year:CountryGermany","Estimate"]
israel <- mdl_m["Year:CountryIsrael","Estimate"]
japan <- mdl_m["Year:CountryJapan","Estimate"]
uk <- mdl_m["Year:CountryUnited Kingdom","Estimate"]
us <- mdl_m["Year:CountryUnited States","Estimate"]



# Adds estimate of each continent to year estimate
exp_est <- data.frame(c("Germany", "Israel", "Japan", "United Kingdom", "United States"),
                            c(yr+germany, yr+israel, yr+japan, yr+uk, yr+us))


names(exp_est) <- c("Country", "expectancy_per_year")
exp_est
```

